HW1 George W Nakhla


Package installs:
```{r}
#install.packages("MPV")
library("MPV")
library("knitr")
```

Q1: Question 2.3. Table B.2 presents data collected during a solar energy project at Georgia
Tech

a) Fit a simple linear regression model relating total heat flux y (kilowatts) to the radial
deflection of the deflected rays x4 (milliradians).
```{r}
solarData <- table.b2
fit <- lm( y ~ x4, data = solarData)
lm( y ~ x4, data = solarData)
plot(solarData$x4, solarData$y, 
     xlab = "radial deflection of deflected rays (x4)", 
     ylab = "total heat flux (y)",
     main = "Model")
abline(fit)
```


So we have the equation y = 607.1 -21.4(x4) as plotted above


b) Construct the analysis-of-variance table and test for significance of regression.
```{r}
summary(fit)
anova_result <- anova(fit)
print(anova_result)
```

72.05% of the variance in y is explained by x4
<br>
The p value of the F statistic is extremely low, indicating that the overall model is significant
<br>

c) Find a 99% CI on the slope.
```{r}
conf_interval <- confint(fit, level = 0.99)
print(conf_interval)

```
After running this and seeing the results, we can say that we are 99% certain that the true slope of the regression (beta1) falls
between:
<br>
(-28.50995, -14.29497)

d) Calculate R^2
```{r}
#previously printed the summary, but this will show only r^2
summary(fit)$r.squared
```
Again, we can say that 72.05% of the variance in y is explained by x4 based on this r^2 value
e) Find a 95% CI on the mean heat flux when the radial deflection is 16.5 milliradians.

```{r}
# Define the new data point
new_data <- data.frame(x4 = 16.5)

# Predict the mean heat flux and obtain confidence intervals
prediction <- predict(fit, newdata = new_data, interval = "confidence", level = 0.95)

# Print the prediction and confidence intervals
print(prediction)
```

We see that based on the model, we get an expected heat flux of 253.9627 kilowatts. We are 95% certain that the true expected heat flux would be between:
<br>
(249.1468, 258.7787)

